---
title: "Native tree sequences in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Native tree sequences in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`Rtskit` keeps a succinct tree sequence in an owned native
`tsk_treeseq_t`. R receives copied summaries or table columns only when an
accessor requests them. No Python runtime or system tskit shared library is
involved.

```{r}
library(Rtskit)

trees <- tskit_example()
trees
tskit_summary(trees)
```

## Tables and coordinates

Node, edge, and tree IDs preserve tskit's zero-based indexing. Genomic
intervals are half-open: an edge over `[left, right)` includes `left` and
excludes `right`.

```{r}
tskit_nodes(trees)
tskit_edges(trees)
tskit_trees(trees)
```

## Round trips

`tskit_dump()` writes the native value directly through the tskit C API.
`tskit_load()` creates a new, independently owned native value.

```{r}
path <- tempfile(fileext = ".trees")
tskit_dump(trees, path)
copy <- tskit_load(path)
identical(tskit_summary(copy), tskit_summary(trees))
unlink(path)
```

The package deliberately does not reinterpret arbitrary table metadata.
Metadata schemas and bytes require explicit schema-aware handling rather than
implicit JSON conversion.
